Evolution of pre-copulatory and post-copulatory inbreeding behaviour

### 2016-10-25 pre_post_biorxiv.tex
Initial biorXiv format; still must add three figures accurately

---
title: "Evolution of pre-copulatory and post-copulatory inbreeding behaviour"
author: "A. Bradley Duthie et al."
date: "August 24, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 2
csl: evolution.csl
bibliography: ms_pre_post.bib
---

******************************

Abstract
===================================================

Inbreeding avoidance has been widely observed through pre-copulatory and post-copulatory mechanisms as a consequence of inbreeding depression in offspring viability. However, no conceptual framework compares evolution of pre-copulatory and post-copulatory inbreeding avoidance jointly, and given biologically relevant costs of inbreeding strategies. Critically, evolution of post-copulatory inbreeding avoidance is restricted to females that mate multiply, potentially leading to complex evolutionary dynamics among inbreeding strategies and polyandry. Accordingly, we use individual-based modelling to track evolution of polyandry and inbreeding strategies under multiple costs and constraints to address three questions: Will evolution of post-copulatory inbreeding avoidance strengthen selection for polyandry? Will selection for a costly inbreeding strategy be negligible given a non-costly alternative strategy? And will selection for an inbreeding strategy be negligible given existence of an already adaptive alternative strategy? We show that selection for polyandry is greatly strengthened by post-copulatory inbreeding avoidance. Strong selection for relatively low cost inbreeding avoidance strategies was observed, while selection for functionally redundant high cost strategies decreased over generations. Finally, existence of an adaptive inbreeding strategy precluded evolution of post-copulatory, but not pre-copulatory, inbreeding avoidance. Our model thereby introduces a novel framework for predicting evolution of both pre-copulatory and post-copulatory inbreeding strategies in empirical systems.

Introduction
===================================================

Inbreeding typically reduces the fitness of inbred offspring (termed 'inbreeding depression'), potentially driving evolution of inbreeding avoidance when inbreeding depression is strong [@Parker1979; @Parker2006; @Pusey1996; @Charlesworth1999; @Keller2002; @Charlesworth2009; @Szulkin2012]. For focal females, inbreeding avoidance might be acheived by avoiding related potential mates (pre-copulatory inbreeding avoidance), or by biasing fertilisation after mating has occurred so that related mates are less likely to sire offspring (post-copulatory inbreeding avoidance). Importantly, evolution of post-copulatory inbreeding avoidance requires that females be polyandrous, mating with multiple males during a single reproductive bout to ensure that post-copulatory choice can be expressed. As a reproductive strategy, polyandry thereby simultaneously allows females to choose multiple mates and creates the opportunity for further cryptic choice enacted through post-copulatory mechanisms. Polyandry has consequently been widely hypothesised to evolve as an adaptation to allow females to avoid inbreeding and therefore produce higher fitness offspring [@Zeh1997; @Jennions2000; @Tregenza2002; @Akcay2007]. However, how pre- and post-copulatory strategies of inbreeding avoidance evolve jointly, and how such evolution is predicted to feed back to affect evolution of polyandry, still remains largely unknown. While pre-copulatory inbreeding avoidance can facilitate evolution of polyandry under some restrictive conditions [@Duthie], the opportunity for post-copulatory choice might facilitate evolution of polyandry under a much wider range of conditions by allowing females to bias fertilisation toward non-kin.

Severe inbreeding depression might cause selection for alleles underlying active pre- and post-copulatory inbreeding avoidance and associated polyandry, but if evolution of one strategy renders its complement redundant for avoiding inbreeding, then selection might instead be sustained over generations for only one strategy. Empirical studies of multiple species have found evidence for female pre-copulatory inbreeding avoidance [e.g., @Lihoreau2007; @Kuriwada2011; @Kingma2013; @Fischer2015], and for post-copulatory inbreeding avoidance [e.g., @Simmons2006; @Bretman2009]. Further, evidence of both pre-copulatory and post-copulatory inbreeding avoidance has been found for a handful of well-studied species, including Trinidadian guppies [*Poecilia reticulata*; @Gasparini2011; @Daniel2015] and house mice [*Mus domesticus*; @Potts1991; @Firman2015], while @Liu2014 found evidence for pre-copulatory, but not post-copulatory, inbreeding avoidance in an experimental study of cabbage beetles (*Colaphellus bowringi*). Other studies do not find any evidence for inbreeding avoidance despite strong inbreeding depression, suggesting that evolution of pre- or post-copulatory inbreeding avoidance cannot be presumed [e.g., @Frere2010; @Reid2014; @Reid2015]. These empirical studies demonstrate that all combinations of adaptive pre-copulatory and post-copulatory inbreeding avoidance are possible, but no conceptual framework exists to predict *a priori* what combinations should be favoured by selection when both pre- and post-copulatory inbreeding avoidance, and polyandry, can evolve.

Frequencies of alleles underlying active pre- and post-copulatory inbreeding avoidance and polyandry are likely to be affected by direct negative selection on phenotypes (i.e., costs). For example, polyandrous females often pay a cost associated with searching or making themselves available to mates, thereby increasing their own risk of predation [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]. Females that express pre-copulatory choice (e.g., reluctance to mate with relatives) might pay a cost of increased risk of mating failure, or harm caused by sexual conflict [@Rowe1994; @Kokko2013]. Females that express post-copulatory choice might pay an energetic cost associated with physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b]. If the relative costs affecting pre- and post-copulatory inbreeding avoidance differ, then alleles underlying the less costly inbreeding strategy might become fixed in the population over generations, while alleles underlying the more costly inbreeding strategy might become extinct if their effects are both costly and made redundant by the less costly strategy.

In addition to the relative importance of costs, evolution of pre- or post-copulatory inbreeding avoidance might be precluded if inbreeding avoidance is already acheived through a complementary strategy, thereby reducing net benefits. Previous models considered the fate of a rare allele underlying inbreeding avoidance in a population initially fixed for random mating [e.g., Duthie and Reid *In press*; @Parker1979; @Parker2006; @Duthie]. Such models thereby isolate the invasion fitness of a single focal strategy. However, when multiple distinct strategies affect realised inbreeding, it cannot be assumed that both will invade simultaneously, nor that invasion fitness of one inbreeding strategy will be independent of another. For example, if pre-adaptation or rapid selection results in the fixation of alleles underlying adaptive pre-copulatory inbreeding avoidance, then alleles underlying post-copulatory inbreeding avoidance might be precluded from invading a population in which post-copulatory choice is not already expressed because the effect that such invading alleles have on decreasing total inbreeding avoidance could be weak or negligible.

In a recent model, @Duthie showed that selection for alleles underlying polyandry and pre-copulatory inbreeding avoidance occurred only under highly constrained conditions in which direct costs of polyandry were low, and when very few males were available during a female's initial mate choice but many males were available during a female's additional mate choice(s), or when polyandry was conditionally expressed based on the degree to which a focal female engaged in inbreeding given her initially chosen mate. In the absence of these constraints, expression of polyandry was expected to increase inbreeding with each additional mate because remaining males were more likely to be relatives once available non-relatives were chosen. However, @Duthie assumed that polyandrous females had no control over the assignment of paternity to their offspring. If after engaging in polyandry, females could bias paternity through post-copulatory mechanisms, then selection for polyandry to avoid inbreeding might be more widely predicted when inbreeding depression is severe [@Zeh1997], potentially feeding back to drive further evolution of pre- and post-copulatory strategies.

@@

******************************

Model
===================================================

We model evolution of polyandry and inbreeding avoidance through pre- and post-copulatory mechanisms by tracking interactions among discrete individuals in a small focal population that persists over multiple generations (key individual traits, parameter values, and variables are described in Table 1). Individuals within this population are either female or male, and can be related to each other if they share a common ancestor. Each individual has 10 diploid loci and therefore 20 alleles underlying each trait: tendency to engage in polyandry ($P_{a}$), pre-copulatory inbreeding avoidance ($M_{a}$, i.e., 'mating' alleles), and post-copulatory inbreeding avoidance ($F_{a}$ i.e., 'fertilisation' alleles). Alleles combine additively to affect genotypes ($P_{g}$, $M_{g}$, and $F_{g}$) and phenotypes ($P_{p}$, $M_{p}$, and $F_{p}$) for each trait. Over generations, we record mean values of alleles underlying polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy. One generation proceeds with females paying costs, expressing polyandry, mating, and fertilising offspring; offspring alleles then mutate and offspring express inbreeding depression; finally, immigrants arrive in the populatoin and density regulation limits population growth.

Costs
----------------------------------------------

@@

Polyandry
----------------------------------------------

@@

The value of $P_{p}$ is affected by the summed values of a focal female's 20 $P_{a}$ alleles. Inheritence of all alleles models a diploid genetic architecture in which 10 $P_{a}$ alleles are randomly sampled from each parent with no physical linkage. Values of $P_{a}$ can take any real number [continuum-of-alleles model; @Kimura1965; @Lande1976; @Reeve2000; @Bocedi2014]. The tendency for polyandry genotype $P_{g}$ is therefore defined by the sum of all of a focal female's 20 $P_{a}$ values, which might be negative or positive. However, the tendency for polyandry phenotype $P_{p}$ is restricted to zero or positive values because the number of males that a female chooses to mate with cannot be negative. Hence, negative values of $P_{g}$ are translated to $P_{p}$ values of zero such that $P_{p} = 0$ when $P_{g} < 0$, but when $P_{g} \geq 0$, then $P_{p} = P_{g}$ (i.e., the phenotype value of polyandry equals the genotype value of polyandry if and only if the genotype value is greater than or equal to zero, else the phenotype value of polyandry always equals zero). The resulting $P_{p}$ is therefore modelled as a threshold trait [@Lynch1998; @Roff1996; @Roff1998; @Duthie], in which polyandry is expressed discretely at the threshold value of $P_{g} > 0$, but is determined by the consequence of continuous genetic variation arising from the effects of $P_{a}$ alleles that each contribute additively to polyandry. Negative genotype values of $P_{g}$ therefore all have the same phenotype, resulting in consistently monandrous females. In contrast, positive values of $P_{g}$ and therefore $P_{p}$ can lead to different degrees of realised polyandry in the form of discrete numbers of mates ($N_{males}$).

Mating
----------------------------------------------

Each male within the population is a potential mate for any female to choose. We therefore assume that there are no opportunity costs associated with male mating such that mating with one female would reduce a male's ability to mate with and sire the offspring of any other female [e.g., @Waser1986]. While all males are therefore available as potential mates, we assume that a female mates with her total allotment of $N_{males}$ without replacement, meaning that $N_{males}$ models the total number of distinct males that a female chooses to mate with rather than the number of times that she mates. In the unlikely event that a female's $N_{males}$ exceeds the total number of males in the population, then she mates with all of the males in the population. Most often, $N_{males}$ will be a small subset of the available males in the population [@Duthie], so the male(s) a focal female chooses will reflect her expression of pre-copulatory inbreeding strategy phenotype ($M_{p}$). Like polyandry allele values, pre-copulatory inbreeding strategy allele values ($M_{a}$) can take any real number. However, unlike polyandry, the sum of all of a female's 20 $M_{a}$ alleles equals both her pre-copulatory inbreeding strategy genotype ($M_{g}$) and her pre-copulatory inbreeding strategy phenotype ($M_{p}$). Negative or positive values of $M_{p}$ cause a female to avoid or prefer mating with kin, respectively. Values of $M_{p}=0$ cause females to mate randomly with respect to kinship.

@@

Fertilisation
----------------------------------------------

@@

Mutation
----------------------------------------------

Alleles of a focal female's offspring mutate with a probability of $\mu=0.001$, with mutation being an independent event that does or does not occur for each locus. When a mutation occurs at a locus, a new value is sampled from a random normal distribution with a mean of zero and a standard devation of mutation effect size $\mu_{SD}$, $\mathcal{N}(0, \mu_{SD})$. This new value is then added to the original value of the allele that mutates [@Kimura1965; @Lande1976; @Bocedi2014; @Duthie]. The standard deviation of the mutation effect size ($\mu_{SD}$) is set to $\sqrt{1/20}$, which scales the variance of mutation effect size to the total number of alleles affecting $P_{p}$, $M_{p}$, and $F_{p}$ phenotypes [see @Kimura1965; @Lande1976; @Bocedi2014].

Inbreeding depression
----------------------------------------------

>>>>

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in,marginparwidth=2in]{geometry}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{datetime}
\setkomafont{\rmfamily\bfseries\boldmath}
\usepackage{wrapfig,floatrow}
\usepackage{float}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{mathabx}
\usepackage{color}
\usepackage{wasysym}
\usepackage{sidecap}
%\usepackage{caption}
\usepackage{xargs,caption,changepage,ifthen}
\usepackage[demo]{graphicx}

% use Unicode characters - try changing the option if you run into troubles with special characters (e.g. umlauts)
\usepackage[utf8]{inputenc}

% clean citations
%\usepackage{cite}

% hyperref makes references clicky. use \url{www.example.com} or \href{www.example.com}{description} to add a clicky url
\usepackage{nameref,hyperref}
\hypersetup{
     colorlinks   = true,
     citecolor    = blue,
     linkcolor    = black
}

% line numbers
\usepackage[right]{lineno}

% improves typesetting in LaTeX
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% text layout - change as needed
%\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Remove % for double line spacing
%\usepackage{setspace} 
%\doublespacing

% use adjustwidth environment to exceed text width (see examples in text)
\usepackage{changepage}

% adjust caption style
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,singlelinecheck=off]{caption}

% remove brackets from references
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% headrule, footrule and page numbers
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}

% use \textcolor{color}{text} for colored text (e.g. highlight to-do areas)
\usepackage{color}

% define custom colors (this one is for figure captions)
\definecolor{Gray}{gray}{.25}

% this is required to include graphics
\usepackage{graphicx}

% use if you want to put caption to the side of the figure - see example in text
\usepackage{sidecap}

% use for have text wrap around figures
\usepackage{wrapfig}
\usepackage[pscoord]{eso-pic}
\usepackage[fulladjust]{marginnote}
\reversemarginpar

%----------------------------------------
%COMMAND FOR DOING SIDE CAPTION.
%----------------------------------------
\newcommandx{\mycaptionminipage}[3][3=c,usedefault]{%
    \begin{minipage}[#3]{#1}%
        \ifthenelse{\equal{#3}{b}}{\captionsetup{aboveskip=0pt}}{}
        \ifthenelse{\equal{#3}{t}}{\captionsetup{belowskip=0pt}}{}
        \vspace{0pt}\centering\captionsetup{width=\textwidth} %Temporarily set caption width
        #2%
    \end{minipage}%
}%
\newcommandx{\mysidecaption}[4][4=c,usedefault]{%
    \checkoddpage%
    \ifoddpage%
        %CASE ODD PAGES
        \mycaptionminipage{\dimexpr\linewidth-#1\linewidth-\intextsep\relax}{#3}[#4]%
        \hfill%
        \mycaptionminipage{#1\linewidth}{#2}[#4]%
    \else%
        %CASE EVEN PAGES
        \mycaptionminipage{#1\linewidth}{#2}[#4]%
        \hfill%
        \mycaptionminipage{\dimexpr\linewidth-#1\linewidth-\intextsep\relax}{#3}[#4]%
    \fi%
}%

% document begins here
\begin{document}
\vspace*{0.35in}

% title goes here:
\begin{flushleft}
{\LARGE
\textbf\newline{Evolution of pre-copulatory and post-copulatory inbreeding behaviour}
}
\newline
% authors go here:
\\
A. Bradley Duthie\textsuperscript{1,*},
Jane M. Reid\textsuperscript{1},
\\
\bigskip
\bf{1} Institute of Biological and Environmental Sciences, School of Biological Sciences, Zoology Building, Tillydrone Avenue, University of Aberdeen, Aberdeen AB24 2TZ, United Kingdom
%\\
%\bf{2} Affiliation B
\\
\bigskip
*  aduthie@abdn.ac.uk, brad.duthie@gmail.com

\end{flushleft}

\section*{Abstract}
\marginpar{
\vspace{0.0cm} % adjust vertical position relative to text with \vspace{} - note that you can enter negative numbers to move margin caption up
\color{Gray} % this gives caption a grey color to set it apart from text body
\textbf{Key words:} % note that \ref{fig1} refers to the corresponding wrapfigure
Inbreeding, inbreeding avoidance, inbreeding depression, mate choice, relatedness, reproductive strategy
}
Inbreeding avoidance has been widely observed through pre-copulatory and post-copulatory mechanisms as a consequence of inbreeding depression in offspring viability. However, no conceptual framework compares evolution of pre-copulatory and post-copulatory inbreeding avoidance jointly, and given biologically relevant costs of inbreeding strategies. Critically, evolution of post-copulatory inbreeding avoidance is restricted to females that mate multiply, potentially leading to complex evolutionary dynamics among inbreeding strategies and polyandry. Accordingly, we use individual-based modelling to track evolution of polyandry and inbreeding strategies under multiple costs and constraints to address three questions: Will evolution of post-copulatory inbreeding avoidance strengthen selection for polyandry? Will selection for a costly inbreeding strategy be negligible given a non-costly alternative strategy? And will selection for an inbreeding strategy be negligible given existence of an already adaptive alternative strategy? We show that selection for polyandry is greatly strengthened by post-copulatory inbreeding avoidance. Strong selection for relatively low cost inbreeding avoidance strategies was observed, while selection for functionally redundant high cost strategies decreased over generations. Finally, existence of an adaptive inbreeding strategy precluded evolution of post-copulatory, but not pre-copulatory, inbreeding avoidance. Our model thereby introduces a novel framework for predicting evolution of both pre-copulatory and post-copulatory inbreeding strategies in empirical systems.

\section*{Introduction}

Inbreeding typically reduces the fitness of inbred offspring (termed `inbreeding depression'), potentially driving evolution of inbreeding avoidance when inbreeding depression is strong \cite[][]{Parker1979, Parker2006, Pusey1996, Charlesworth1999, Keller2002, Charlesworth2009, Szulkin2012}. For focal females, inbreeding avoidance might be acheived by avoiding related potential mates (pre-copulatory inbreeding avoidance), or by biasing fertilisation after mating has occurred so that related mates are less likely to sire offspring (post-copulatory inbreeding avoidance). Importantly, evolution of post-copulatory inbreeding avoidance requires that females be polyandrous, mating with multiple males during a single reproductive bout to ensure that post-copulatory choice can be expressed. As a reproductive strategy, polyandry thereby simultaneously allows females to choose multiple mates and creates the opportunity for further cryptic choice enacted through post-copulatory mechanisms. Polyandry has consequently been widely hypothesised to evolve as an adaptation to allow females to avoid inbreeding and therefore produce higher fitness offspring \cite[][]{Zeh1997, Jennions2000, Tregenza2002, Akcay2007}. However, how pre- and post-copulatory strategies of inbreeding avoidance evolve jointly, and how such evolution is predicted to feed back to affect evolution of polyandry, still remains largely unknown. While pre-copulatory inbreeding avoidance can facilitate evolution of polyandry under some restrictive conditions \cite[][]{Duthie}, the opportunity for post-copulatory choice might facilitate evolution of polyandry under a much wider range of conditions by allowing females to bias fertilisation toward non-kin.

Severe inbreeding depression might cause selection for alleles underlying active pre- and post-copulatory inbreeding avoidance and associated polyandry, but if evolution of one strategy renders its complement redundant for avoiding inbreeding, then selection might instead be sustained over generations for only one strategy. Empirical studies of multiple species have found evidence for female pre-copulatory inbreeding avoidance \cite[e.g.,][]{Lihoreau2007, Kuriwada2011, Kingma2013, Fischer2015}, and for post-copulatory inbreeding avoidance \cite[e.g.,][]{Simmons2006, Bretman2009}. Further, evidence of both pre-copulatory and post-copulatory inbreeding avoidance has been found for a handful of well-studied species, including Trinidadian guppies \cite[\textit{Poecilia reticulata};][]{Gasparini2011, Daniel2015} and house mice \cite[\textit{Mus domesticus};][]{Potts1991, Firman2015}, while \cite{Liu2014} found evidence for pre-copulatory, but not post-copulatory, inbreeding avoidance in an experimental study of cabbage beetles (\textit{Colaphellus bowringi}). Other studies do not find any evidence for inbreeding avoidance despite strong inbreeding depression, suggesting that evolution of pre- or post-copulatory inbreeding avoidance cannot be presumed \cite[e.g.,][]{Frere2010, Reid2014, Reid2015}. These empirical studies demonstrate that all combinations of adaptive pre-copulatory and post-copulatory inbreeding avoidance are possible, but no conceptual framework exists to predict \textit{a priori} what combinations should be favoured by selection when both pre- and post-copulatory inbreeding avoidance, and polyandry, can evolve.

Frequencies of alleles underlying active pre- and post-copulatory inbreeding avoidance and polyandry are likely to be affected by direct negative selection on phenotypes (i.e., costs). For example, polyandrous females often pay a cost associated with searching or making themselves available to mates, thereby increasing their own risk of predation \cite[e.g.,][]{Rowe1988, Ronkainen1994, Koga1998}. Females that express pre-copulatory choice (e.g., reluctance to mate with relatives) might pay a cost of increased risk of mating failure, or harm caused by sexual conflict \cite[][]{Rowe1994, Kokko2013}. Females that express post-copulatory choice might pay an energetic cost associated with physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation \cite[e.g.,][]{Gasparini2011, Tuni2013, Fitzpatrick2014b}. If the relative costs affecting pre- and post-copulatory inbreeding avoidance differ, then alleles underlying the less costly inbreeding strategy might become fixed in the population over generations, while alleles underlying the more costly inbreeding strategy might become extinct if their effects are both costly and made redundant by the less costly strategy.

In addition to the relative importance of costs, evolution of pre- or post-copulatory inbreeding avoidance might be precluded if inbreeding avoidance is already acheived through a complementary strategy, thereby reducing net benefits. Previous models considered the fate of a rare allele underlying inbreeding avoidance in a population initially fixed for random mating \cite[e.g.,][; Duthie and Reid \textit{In press}]{Parker1979, Parker2006, Duthie}. Such models thereby isolate the invasion fitness of a single focal strategy. However, when multiple distinct strategies affect realised inbreeding, it cannot be assumed that both will invade simultaneously, nor that invasion fitness of one inbreeding strategy will be independent of another. For example, if pre-adaptation or rapid selection results in the fixation of alleles underlying adaptive pre-copulatory inbreeding avoidance, then alleles underlying post-copulatory inbreeding avoidance might be precluded from invading a population in which post-copulatory choice is not already expressed because the effect that such invading alleles have on decreasing total inbreeding avoidance could be weak or negligible.

In a recent model, \cite{Duthie} showed that selection for alleles underlying polyandry and pre-copulatory inbreeding avoidance occurred only under highly constrained conditions in which direct costs of polyandry were low, and when very few males were available during a female's initial mate choice but many males were available during a female's additional mate choice(s), or when polyandry was conditionally expressed based on the degree to which a focal female engaged in inbreeding given her initially chosen mate. In the absence of these constraints, expression of polyandry was expected to increase inbreeding with each additional mate because remaining males were more likely to be relatives once available non-relatives were chosen. However, \cite{Duthie} assumed that polyandrous females had no control over the assignment of paternity to their offspring. If after engaging in polyandry, females could bias paternity through post-copulatory mechanisms, then selection for polyandry to avoid inbreeding might be more widely predicted when inbreeding depression is severe \cite[][]{Zeh1997}, potentially feeding back to drive further evolution of pre- and post-copulatory strategies.

@@

\section*{Model}

\begin{table}[!ht]
\begin{adjustwidth}{-1.75in}{0in}
\centering
\caption{\color{Gray}Individual traits (A), model parameter values (B), and model variables (C) for an individual based model of the evolution of polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy.}
\begin{tabular}{lllll}
\hline
A & Trait & Allele & Genotype & Phenotype &
\hline
  & Tendency for polyandry               &   $P_{a}$  &  $P_{g}$  &  $P_{p}$  &
  & Pre-copulatory inbreeding strategy   &   $M_{a}$  &  $M_{g}$  &  $M_{p}$  &
  & Post-copulatory inbreeding strategy  &   $F_{a}$  &  $F_{g}$  &  $F_{p}$  &
  &                                      &            &           &           &
\hline
B & Description & Parameter & & Default value(s) &
\hline

  & Cost of pre-copulatory inbreeding strategy    & $c_{M}$    & & $0$, $0.02$   &
  & Cost of post-copulatory inbreeding strategy   & $c_{F}$    & & $0$, $0.02$   &
  & Focal female's number of offspring            & $n$        & & $8$           &
  & Log-linear slope of inbreeding depression     & $\beta$    & & $3$           &
  & Adult male immigrants per generation          & $\rho$     & & $5$           & 
  & Female carrying capacity                      & $K_{f}$    & & $100$         &
  & Male carrying capacity                        & $K_{m}$    & & $100$         &
  & Mutation rate of alleles                      & $\mu$      & & $0.001$       &
  & Standard deviation of mutation effect size    & $\mu_{SD}$ & & $\sqrt{1/20}$ &
  &                                               &            & &               &
\hline
C & Description & & & Variable &
\hline
  & Coefficient of kinship                                            & & & $k$              &
  & Focal female's number of mates                                    & & & $N_{males}$      &
  & Female $i$'s perceived absolute mate quality of male $j$          & & & $Q^{m}_{ij}$     &
  & Female $i$'s perceived relative mate quality of male $j$          & & & $q^{m}_{ij}$     &
  & Female $i$'s perceived absolute fertilisation quality of male $j$ & & & $Q^{f}_{ij}$     &
  & Female $i$'s perceived relative fertilisation quality of male $j$ & & & $q^{f}_{ij}$     &
  & Viability of a focal female's offspring                           & & & $\Psi_{off}$     &
\hline
\end{tabular}
\end{adjustwidth}
\end{table}

We model evolution of polyandry and inbreeding avoidance through pre- and post-copulatory mechanisms by tracking interactions among discrete individuals in a small focal population that persists over multiple generations (key individual traits, parameter values, and variables are described in Table 1). Individuals within this population are either female or male, and can be related to each other if they share a common ancestor. Each individual has 10 diploid loci and therefore 20 alleles underlying each trait: tendency to engage in polyandry ($P_{a}$), pre-copulatory inbreeding avoidance ($M_{a}$, i.e., 'mating' alleles), and post-copulatory inbreeding avoidance ($F_{a}$ i.e., `fertilisation' alleles). Alleles combine additively to affect genotypes ($P_{g}$, $M_{g}$, and $F_{g}$) and phenotypes ($P_{p}$, $M_{p}$, and $F_{p}$) for each trait. Over generations, we record mean values of alleles underlying polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy. One generation proceeds with females paying costs, expressing polyandry, mating, and fertilising offspring; offspring alleles then mutate and offspring express inbreeding depression; finally, immigrants arrive in the populatoin and density regulation limits population growth.

\subsection*{Costs}

@@

\subsection*{Polyandry}

@@

The value of $P_{p}$ is affected by the summed values of a focal female's 20 $P_{a}$ alleles. Inheritence of all alleles models a diploid genetic architecture in which 10 $P_{a}$ alleles are randomly sampled from each parent with no physical linkage. Values of $P_{a}$ can take any real number \cite[continuum-of-alleles model;][]{Kimura1965, Lande1976, Reeve2000, Bocedi2014}. The tendency for polyandry genotype $P_{g}$ is therefore defined by the sum of all of a focal female's 20 $P_{a}$ values, which might be negative or positive. However, the tendency for polyandry phenotype $P_{p}$ is restricted to zero or positive values because the number of males that a female chooses to mate with cannot be negative. Hence, negative values of $P_{g}$ are translated to $P_{p}$ values of zero such that $P_{p} = 0$ when $P_{g} < 0$, but when $P_{g} \geq 0$, then $P_{p} = P_{g}$ (i.e., the phenotype value of polyandry equals the genotype value of polyandry if and only if the genotype value is greater than or equal to zero, else the phenotype value of polyandry always equals zero). The resulting $P_{p}$ is therefore modelled as a threshold trait \cite[][]{Lynch1998, Roff1996, Roff1998, Duthie}, in which polyandry is expressed discretely at the threshold value of $P_{g} > 0$, but is determined by the consequence of continuous genetic variation arising from the effects of $P_{a}$ alleles that each contribute additively to polyandry. Negative genotype values of $P_{g}$ therefore all have the same phenotype, resulting in consistently monandrous females. In contrast, positive values of $P_{g}$ and therefore $P_{p}$ can lead to different degrees of realised polyandry in the form of discrete numbers of mates ($N_{males}$).

\subsection*{Mating}

Each male within the population is a potential mate for any female to choose. We therefore assume that there are no opportunity costs associated with male mating such that mating with one female would reduce a male's ability to mate with and sire the offspring of any other female \cite[e.g.,][]{Waser1986}. While all males are therefore available as potential mates, we assume that a female mates with her total allotment of $N_{males}$ without replacement, meaning that $N_{males}$ models the total number of distinct males that a female chooses to mate with rather than the number of times that she mates. In the unlikely event that a female's $N_{males}$ exceeds the total number of males in the population, then she mates with all of the males in the population. Most often, $N_{males}$ will be a small subset of the available males in the population \cite[][]{Duthie}, so the male(s) a focal female chooses will reflect her expression of pre-copulatory inbreeding strategy phenotype ($M_{p}$). Like polyandry allele values, pre-copulatory inbreeding strategy allele values ($M_{a}$) can take any real number. However, unlike polyandry, the sum of all of a female's 20 $M_{a}$ alleles equals both her pre-copulatory inbreeding strategy genotype ($M_{g}$) and her pre-copulatory inbreeding strategy phenotype ($M_{p}$). Negative or positive values of $M_{p}$ cause a female to avoid or prefer mating with kin, respectively. Values of $M_{p}=0$ cause females to mate randomly with respect to kinship.

@@

\subsection*{Fertilisation}

@@

\subsection*{Mutation}

Alleles of a focal female's offspring mutate with a probability of $\mu=0.001$, with mutation being an independent event that does or does not occur for each locus. When a mutation occurs at a locus, a new value is sampled from a random normal distribution with a mean of zero and a standard devation of mutation effect size $\mu_{SD}$, $\mathcal{N}(0, \mu_{SD})$. This new value is then added to the original value of the allele that mutates \cite[][]{Kimura1965, Lande1976, Bocedi2014, Duthie}. The standard deviation of the mutation effect size ($\mu_{SD}$) is set to $\sqrt{1/20}$, which scales the variance of mutation effect size to the total number of alleles affecting $P_{p}$, $M_{p}$, and $F_{p}$ phenotypes \cite[see][]{Kimura1965, Lande1976, Bocedi2014}.

\subsection*{Inbreeding depression}

The viability of a focal female $i$'s offspring ($\Psi_{\textrm{off}}$) decreases as a function of the coefficient of kinship between $i$ and the sire of her offspring $j$ ($k_{ij}$), and a log-linear slope of inbreeding depression $\beta$. We model selection as having an absolute effect on offspring viability (i.e., hard selection), rather than a relative effect on viability (i.e., soft selection), so that the effect of $\beta$ is consistent over generations and across different parameter combinations. Natural selection on tendency for polyandry ($P_{p}$), pre-copulatory inbreeding strategy ($M_{p}$), and post-copulatory inbreeding strategy ($F_{p}$) can therefore be compared across generations and parameter combinations. If offspring viability was instead relative such that the least inbred offspring had a high viability regardless of the absolute value of $k_{ij}$, then the effect of kinship coefficients and inbreeding depression slopes on offspring viability would depend on the population-wide distribution of $k_{ij}$ between females and males, making model inference more difficult. Therefore, we define offspring viability as,
\begin{equation}
\Psi_{\textrm{off}} = e^{-\beta k_{ij}}
\end{equation}
The slope $\beta$ models the number of haploid lethal equivalents that exist as deleterious recessive alleles in the gametes of $i$ and $j$, and which might combine to become homozygous in offspring and reduce viability. Equation 1 assumes that each deleterious recessive allele affects offspring viability independently, meaning that every deleterious recessive allele added will have a multiplicative affect on decreasing total viability \cite[][]{Morton1956, Mills1994}, causing the log-linear relationship between parent kinship and inbreeding depression. We assume that inbreeding depression always decreases offspring viability, $\beta > 0$ (i.e., no outbreeding depression). Therefore, because $k_{ij}$ is constrained to values between zero and one, the product of $-\beta$ and $k_{ij}$ must be a value less than or equal to zero, $-\beta \times k_{ij} \leq 0$. Values of $\Psi_{\textrm{off}}$ must therefore be between zero (if $-\beta \times k_{ij}$ is very negative) and one (if $-\beta \times k_{ij} = 0$). We can therefore define $\Psi_{\textrm{off}}$ as the probability that an offspring is viable, and sample its realised viability using a Bernoulli trial.

\subsection*{Immigration}

After offspring viability is realised, $\rho$ adult immigrants are added to the focal population. The coefficient of kinship between an immigrant and all other individuals (including native individuals and other immigrants) always equals zero ($k_{ij}=0$). Because kinship coefficient values are calculated directly from the pedigree of the focal population \cite[not from individuals' genomes; see][]{Duthie}, the addition of immigrants introduces non-kin into the population at a constant rate and prevents the mean kinship coefficient of the population from asymptoting to a value of 1 over generations. If the focal population was instead closed ($\rho=0$), then all alleles would quickly be expected to become identical-by-descent among descendants of the ancestral population. To avoid having immigrants affect evolution of tendency for polyandry ($P_{p}$), pre-copulatory inbreeding strategy ($M_{p}$), and post-copulatory inbreeding strategy ($F_{p}$), immigrants are always male and can therefore be chosen based on their kinship values of zero but not actively affect reproductive decisions through the expression of $P_{p}$, $M_{p}$, or $F_{p}$. To further avoid having immigrants affect evolutionary dynamics, allele values underlying ($P_{a}$), pre-copulatory inbreeding strategy ($M_{a}$), and post-copulatory inbreeding strategy ($F_{a}$) are randomly sampled from normal distributions with means and standard deviations of $P_{a}$, $M_{a}$, and $F_{a}$ values equal to the means and standard deviations of those in the focal population at the time of immigration. Overall, this model of immigration effectively assumes that the focal population receives immigrants from other nearby populations that are subject to the same selection on $P_{p}$, $M_{p}$, and $F_{p}$.

\subsection*{Density regulation}
